Formalization of Typed Lua

### 2015-06-11 doc/papers/paper/paper.tex
more changes to section 2 of the DLS paper, reworked the description of the different table types

Table types represent the various forms
that Lua tables can take.
The syntactical form of table types is $\{ K_{1}{:}V_{1}, ..., K_{n}{:}V_{n} \}_{tag}$,
where each $K_{i}$ represents the type of a table key,
and each $V_{i}$ represents the type of the value that table keys of type $K_{i}$ map to.
Key types can only be literal types, base types, or the top type.
We made this restriction to the type of the keys because the statistics
that we discussed in Section \ref{sec:statistics} showed that most
of the tables are records, lists, and hashes.
The type $\Value$ is an option when we need a loose table type.
For instance, $\{\Value:\Value\}_{closed}$ represents the type of a
table in which both indices and values can have any type.
Value types can be any first-level type, and can optionally include
the $\Const$ type to denote immutable values.

We also use the tags \emph{unique}, \emph{open}, \emph{fixed}, and \emph{closed}
to classify table types.
The tag \emph{unique} represents tables with no keys that do not
inhabit one of the table's key types, and with no alias.
In particular, the type of the table constructor has this tag.
The tag \emph{open} represents \emph{unique} table types that
have at least one alias.
The tag \emph{fixed} represents \emph{unique} table types that we
do not know how many aliases they have.
In particular, the type of a class has this tag.
The tag \emph{closed} represents table types that do not provide
any guarantees about keys with types not listed in the table type.
In particular, in the concrete syntax, type annotations, interface
declarations, and userdata declarations always describe \emph{closed} table types.
In the next sections we explain in more detail why we need
different table types.

Any table type has to be \emph{well-formed}.
Informally, a table type is well-formed if key types do not overlap.
In Section \ref{sec:tables} we formalize the definition of well-formed table types.
We delay the proper formalization of well-formed table types because we use
consistent-subtyping in this formalization.

>>>>

Table types represent some of the forms that Lua
tables can take. What a particular table type denotes
depends on its {\em tag}: the {\em unique}, {\em open}
and {\em fixed} tags all denote tables where it is
guaranteed that every key belongs to one of the {\em key
types} $K_i$. We restrict key types to literal types,
base types, and the top type to make the system simpler
for the user, as our survey of Lua programs showed that
this restriction reflects the way Lua tables are used.
Fields are assignable by default, but can be marked
as non-assignable with the {\bf const} qualifier. Having
a field be {\bf const} is not a guarantee that its
value cannot change, as the table may have aliases with
a non-{\bf const} type for that field.

The difference between {\em unique}, {\em open} and
{\em fixed} tables is the guarantees that the type system
give about the reference to a value of that type. A reference
to an {\em unique} table is guaranteed to have no aliases,
so allows the greater flexibility in reinterpreting its type;
a reference to an {\em open} table is guaranteed to have
only {\em closed} aliases (we will explain the the meaning of the {\em closed} tag shortly), which still lets the type system
reinterpret the type of a reference, but with more
restrictions; a {\em fixed} table can have any number of {\em fixed} or {\em closed} aliases, so its type cannot change
anymore.

Table types with the {\em fixed} tag are suitable for typing tables that are used as maps, lists, or even a mixture of those. The guarantee that there are no keys with a type that is not
one of the key types allows type-safe iteration on these
tables.

This guarantee forbids width subtyping, which is a necessity
for object-oriented programming, so we have the {\em closed}
tag, which removes the guarantee and allows width subtyping
between other table types and {\em closed} table types. A {\em closed} table type is akin to an object-oriented {\em interface}.

Any table type has to be \emph{well-formed}.
Informally, a table type is well-formed if key types do not overlap. A formal definition of this relation depends on
the consistent-subtyping relation which we present in the
next section, so we delay the definition until Section~\ref{sec:tables}, where the typing rules for
table types are explained.
